---
title: "Location quotient LQ. A  Spatial concentration perspective."
author: "JLMR"
date: "2019-04-03"
output: html_document
---


```{r set-global-options, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      echo = FALSE, 
                      cache = FALSE,
                      include = TRUE,
                      collapse = FALSE,
                      dependson = NULL,
                      engine = "R", # Chunks will always have R code, unless noted
                      error = FALSE,                      
                      fig.path="Figuras/",  # Set the figure options
                      fig.align = "center", 
                      fig.width = 7,
                      fig.height = 5)
```



```{r silent-packages, echo = FALSE, eval = TRUE, message=FALSE, include = FALSE}
library(ggthemes)
library(foreign)
library(knitr)
library(tidyverse)
library(DT)
library(plotly)
library(grid)
library(gridExtra)
```


```{r}
library(viridis)
lqtrd<-read.csv("LQ98_17X.csv")

fs<-mutate(lqtrd,Fecha=as.numeric(Fecha),Region=ifelse(Estado=="Coahuila"|Estado== "Nuevo Le칩n"|Estado== "Chihuahua"|Estado== "Baja California"|Estado=="Tamaulipas"|Estado=="Sonora",1,ifelse(Estado=="Chiapas"|Estado=="Yucat치n"|Estado== "Campeche"|Estado== "Tabasco"|Estado=="Quintana Roo",2,0)))%>%
  filter(LQ,Region!=0&Region!=1)

fn<-mutate(lqtrd,Fecha=as.numeric(Fecha),Region=ifelse(Estado=="Coahuila"|Estado== "Nuevo Le칩n"|Estado== "Chihuahua"|Estado== "Baja California"|Estado=="Tamaulipas"|Estado=="Sonora",1,ifelse(Estado=="Chiapas"|Estado=="Yucat치n"|Estado== "Campeche"|Estado== "Tabasco"|Estado=="Quintana Roo",2,0)))%>%
  filter(LQ,Region!=0&Region!=2)


cb_palette <- c("red", "#E69F00", "#56B4E9", "#009E73", "#F0E442","#0072B2", "#D55E00", "#CC79A7")

p1<-ggplot(fs, aes(x=Fecha, y=LQ, group=Estado, colour=Estado))+
geom_line(aes(y=LQ))+
geom_point(size=1.5, shape=21,fill="white")+
geom_hline(yintercept = 1,linetype="dashed")+
xlab("")+ylab("LQ")+
theme_bw()+
labs(colour="State")+
scale_colour_manual(values=cb_palette)

p1

p2<-ggplot(fn, aes(x=Fecha, y=LQ, group=Estado, colour=Estado))+
geom_line(aes(y=LQ))+
geom_point(size=1.5, shape=21,fill="white")+
geom_hline(yintercept = 1,linetype="dashed")+
xlab("")+ylab("LQ")+
theme_bw()+
labs(colour="State")+
scale_colour_viridis_d(option = "plasma")
p2
```

LQ evolution in time, depicted  above, reveals a clearer pattern, an opposite path between the two border regions. While the observed trend suggest an important epidemiological transition towards improved public health at the US-Mexico border, the
opposite holds true for the Mexico-Guatemala-Belize border. 

What social determinants influence this relocation of disease at a regional level?


