---
title: "Location quotient LQ. A  Spatial concentration perspective."
author: "JLMR"
date: "2019-04-03"
output: html_document
---


```{r set-global-options, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      echo = FALSE, 
                      cache = FALSE,
                      include = TRUE,
                      collapse = FALSE,
                      dependson = NULL,
                      engine = "R", # Chunks will always have R code, unless noted
                      error = FALSE,                      
                      fig.path="Figuras/",  # Set the figure options
                      fig.align = "center", 
                      fig.width = 7,
                      fig.height = 5)
```



```{r silent-packages, echo = FALSE, eval = TRUE, message=FALSE, include = FALSE}
library(ggthemes)
library(foreign)
library(knitr)
library(tidyverse)
library(DT)
library(plotly)
library(grid)
library(gridExtra)
```


LQ evolution in time, depicted in above , reveals a clearer pattern, an opposite path between the two border regions. While the observed trend suggest an important epidemiological transition towards improved public health at the US-Mexico border, the
opposite holds true for the Mexico-Guatemala-Belize border. 

What social determinants influence this relocation of disease at a regional level?
