---
title: "Diabetes Mortality growth in México, 20 years of changing public health patterns"
author: "JLMR"
date: "2019-03-29"
output: html_document
---


```{r set-global-options, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      echo = FALSE, 
                      cache = FALSE,
                      include = TRUE,
                      collapse = FALSE,
                      dependson = NULL,
                      engine = "R", # Chunks will always have R code, unless noted
                      error = FALSE,                      
                      fig.path="Figuras/",  # Set the figure options
                      fig.align = "center", 
                      fig.width = 7,
                      fig.height = 5)
```



```{r silent-packages, echo = FALSE, eval = TRUE, message=FALSE, include = FALSE}
library(ggthemes)
library(foreign)
library(knitr)
library(tidyverse)
library(DT)
library(plotly)
library(grid)
library(gridExtra)
```





```{r, echo = FALSE,  fig.width = 10, fig.height = 6  }

library(ggthemes)

growth<-readRDS("tasa_a.rds")%>%
  mutate(fecha=replace(fecha,fecha==99,-1),
         fecha=replace(fecha,fecha==98,-2), 
         fecha=as.numeric(fecha), Diabetes=as.factor(ifelse(Diabetes=="Otras_Causas", "Overall mortality","Diabetes" )))


ave<-growth%>%
filter(fecha==17)%>%
  group_by(d)%>%
  summarize(ave=mean(rate))


p1<-ggplot(growth, aes(x=fecha,  group=Diabetes, color=Diabetes)) +
geom_line(aes(y=rate))+
facet_wrap(~Estado, nrow=6)+
geom_hline(yintercept = 0, linetype="dashed")+
ylab("Growth rate 1998=0 (%)") +
xlab("Year")+
scale_color_manual(values = c('red', "dodgerblue3",""))+
labs(color="", caption ="Source: Own estimation with data from SINAIS 1998-2017", title="Diabetes deaths growth 1998-2017")+
theme(legend.position=c(.8,.02))+
scale_x_continuous(breaks=c(0, 5, 10, 15, 17))+
theme_tufte()+
  theme(legend.position="bottom")

p1
```
<div style="text-align: justify">

During the last fifteen years, T2D mortality showed an explosive pattern at the border
states. This growth rate pattern departs from the trend followed by all other death
causes.


The following graphs shows a closer look a  the trend in each region.</div>


```{r, fig.width = 10, fig.height = 6}

growth<-readRDS("tasa_a.rds")%>%
mutate(fecha=replace(fecha,fecha==99,-1),
fecha=replace(fecha,fecha==98,-2),
fecha=as.numeric(fecha))%>%
select(-c(d,casos))

fn<-data.frame(fn=c("Baja California","Coahuila","Chihuahua","Nuevo León","Sonora", "Tamaulipas"))

fs<-data.frame(fs=c("Campeche","Chiapas","Yucatán","Quintana Roo","Tabasco"))


growth<-mutate(growth,fro=ifelse(Estado%in%fn$fn,1,
                            ifelse(Estado%in%fs$fs,2,0)))


f<-growth%>%
  filter(fro!=0)

p<-ggplot(f, aes(x=fecha,  group=Diabetes, color=Diabetes)) +
geom_line(aes(y=rate))+
facet_wrap(~Estado)+
geom_hline(yintercept = 0, linetype="dashed")+
ylab("Tasa de crecimiento 1998=0 (%)") +
xlab("Año")+
scale_color_manual(values = c('orange', "deepskyblue4",""))+
labs(color="", caption ="Source: Own estimation with data from SINAIS 1998-2017")+
scale_x_continuous(breaks=c(0, 5, 10, 15, 17))+
  theme_classic()+
  labs(title = "Border States 1998-2017")+
  theme(legend.position="bottom")

p

```

<div style="text-align: justify">

In particular, a rapid growth trajectory at the **southern border** is evident, with a widening gap by the end of the study period. This trend reaches a 169% increase in **Quintana Roo** by 2019 relative to the base year 1998, with an annual 8.45% average growth rate.


Nevertheless, the biggest gap between the overall mortality levels and diabetes mortality
is observed in **Chiapas**, with a gap equivalent to **109.8%** between diabetes mortality, which register a **149.8%** accumulated growth during the period 1998-2017 compared to a moderate **39.9%** mortality growth for all other causes. 

Despite the fact  that, among the northern border states there is also an accelerated growth trend,  we find smaller gaps in this region. 

In fact, the smallest gap between Diabetes mortality trends and the overall path is  found in Sonora and it's equivalent to **12.8%**, which is just a reflection of the fact that Diabetes is a main component of  ("pulling") the overall mortality  trend. 

In these states Diabetes is so prevalent that it's behaviour is shaping the over all mortality trayectory.

This contrast relative to the southern border suggest significant and fast regional changes in the Mexican population health, as  well as important implications for the health care system as a whole, it is important to remember that Diabetes is considered a high  financial impact disease from the perspective of the Health care system.   
