---
title: "Crecimiento de la mortalidad por Diabetes en México, 20 años de cambios."
author: "JLMR"
date: "2019-03-29"
output: html_document
---


```{r set-global-options, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      echo = FALSE, 
                      cache = FALSE,
                      include = TRUE,
                      collapse = FALSE,
                      dependson = NULL,
                      engine = "R", # Chunks will always have R code, unless noted
                      error = FALSE,                      
                      fig.path="Figuras/",  # Set the figure options
                      fig.align = "center", 
                      fig.width = 7,
                      fig.height = 5)
```



```{r silent-packages, echo = FALSE, eval = TRUE, message=FALSE, include = FALSE}

library(foreign)
library(knitr)
library(tidyverse)
library(DT)
library(plotly)
library(grid)
library(gridExtra)
```





```{r, echo = FALSE, fig.width = 8, fig.height = 5  }

library(ggthemes)

growth<-readRDS("tasa_a.rds")%>%
  mutate(fecha=replace(fecha,fecha==99,-1),
         fecha=replace(fecha,fecha==98,-2), 
         fecha=as.numeric(fecha))

ave<-growth%>%
filter(fecha==17)%>%
  group_by(d)%>%
  summarize(ave=mean(rate))
169/20

p1<-ggplot(growth, aes(x=fecha,  group=Diabetes, color=Diabetes)) +
geom_line(aes(y=rate))+
facet_wrap(~Estado, nrow=6)+
geom_hline(yintercept = 0, linetype="dashed")+
ylab("Tasa de crecimiento 1998=0 (%)") +
xlab("Año")+
scale_color_manual(values = c('red', "dodgerblue3",""))+
labs(color="")+
theme(legend.position=c(.8,.02))+
scale_x_continuous(breaks=c(0, 5, 10, 15, 17))+
theme_tufte()+
  theme(legend.position="bottom")

p1
```
During the last fifteen years, T2D mortality showed an explosive pattern at the border
states. This growth rate pattern departs from the trend followed by all other death
causes.


In particular, a rapid growth trajectory at the **southern border** is evident, with a widening gap by the end of the study period. This trend reaches a 169% increase in **Quintana Roo** by 2019 relative to the base year 1998, with an annual 8.45% average growth rate.


Nevertheless, the biggest gap between the overall mortality levels and diabetes mortality
is observed in **Chiapas**, with a gap equivalent to **109.8%** between diabetes mortality, which register a **149.8%** accumulated growth during the period 1998-2017 compared to a moderate **39.9%** mortality growth for all other causes. 

Despite the fact  that, among the northern border states there is also an accelerated growth trend,  we find smaller gaps in this region. 

In fact, the smallest gap between Diabetes mortality trends and the overall path is  found in Nuevo León and it's equivalent to **17%**, whhic just reflects the fact that Diabetes is defining "pulling" the observed mortality  trend. In these states Diabetes is so prevalent that it's behaviour is shaping the over all mortality trayectory.

This contrast relative to the southern border suggest significant and fast regional changes in the Mexican population health, as  well as important implications for the health care system as a whole, it is important to remember that Diabetes is considered a high  financial impact disease from the perspective of the Health care system.   

  

